Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-22 ASA2013/asa2013.tex
la2pdf commit - 2013:01:22	19:53:42	(Tue22.Jan)

\usepackage[utf8x]{inputenc}
\usepackage[table,dvipsnames,svgnames]{xcolor}
\usepackage{titlesec}

@@

\ifTODO
   \newcounter{todoidx}
   \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
   \setlength\marginparsep{1pt}
   \setlength\marginparwidth{35pt}
   \newlength\marginparwidthsmall
   \setlength\marginparwidthsmall{\marginparwidth}
   \addtolength\marginparwidthsmall{-7pt}
   \newcommand\todo[1]{%
      \addtocounter{todoidx}{1}%
      {\color{Red}\bf(\thetodoidx{})}%%\fbox{\bf\thetodoidx{}}}%
      \marginpar{%
         {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
         \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{\scriptsize #1}}}}

@@

For even larger systems we suggest employing the LCA beamformer. It does not calculate a weightset, but merely computes the beamformer output for each of a predefined set of weights, and selects the one that best fulfils the MVDR criterion. The LCA creates images with a quality comparable to MVDR, and it is perfectly suited for a GPU.
\end{abstract}
% \ \\
% Total: 195 words (200 max)
% 
% \ \\
% Noen tanker jeg har selv naa:
% \begin{itemize}
% \item Det mangler noe om bruksomraader.
% \item Maa passe paa aa ikke selge LCA saa bra at MVDR ser helt haaplos ut(?)
% \end{itemize}
% 
% 
% \ \\
% Questions they wanted answered:
% \begin{itemize}
% \item This session will feature work on adapting computationally demanding modeling and signal processing tasks to take advantage of various parallel architectures, such as cloud computing, computing clusters, multicore CPUs, multi-core graphic processing units (GPUs), DSP chips, and FPGAs.
% \item If algorithm is slow - What technology will help me?
% \item What skills are needed?
% \item What tools are available and what's the cost?
% \item What acceleration factor are we looking at?
% \item How can I get started?
% \item What we did. Which tools? Was it hard? Speedup?
% \end{itemize}

\section{Introduction}

@@

We propose three different solutions to the complexity issue depending on system size. For smaller systems we use the standard time-domain MVDR method, arithmetically optimize the estimation step, and show that image creation rates of more than 1\,Mpx/s can be achieved when this algorithm is run on a GPU. For medium sized systems we suggest applying the MVDR in the frequency domain (beamspace) instead, which allows for comparable processing rates at a negligible reduction in image quality. Finally, for large systems the estimation and inversion step can be avoided completely by utilising the Low Complexity Adaptive (LCA) beamformer.
% 
% 
% The end design is more than 2 orders of magnitude faster than a C implementation we started off with, it took considerable effort to get there.
% \\
% << MOTIVATION >>
% \\

@@

This article is outlined as follows: In section \ref{methods} we will present background material on the MVDR method operated in time and frequency domain, and on the LCA beamformer. For each method we start with the mathematical formulation, then evaluate complexity issues, and finally give our thoughts on some implementation aspects. The results are collected in \ref{implementation}. Finally, we conclude in \ref{discssion}.

>>>>

% \usepackage[utf8x]{inputenc}
\usepackage[table,dvipsnames,svgnames]{xcolor}
\usepackage{titlesec}

@@

\ifTODO
   \newcounter{todoidx}
   \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
   \setlength\marginparsep{1pt}
   \setlength\marginparwidth{50pt}
   \newlength\marginparwidthsmall
   \setlength\marginparwidthsmall{\marginparwidth}
   \addtolength\marginparwidthsmall{-7pt}
   \newcommand\todo[1]{%
      \addtocounter{todoidx}{1}%
      {\color{Red}\bf(\thetodoidx{})}%%\fbox{\bf\thetodoidx{}}}%
      \marginpar{%
         {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
         \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{\scriptsize #1}}}}

@@

For even larger systems we suggest employing the LCA beamformer. It does not calculate a weightset, but merely computes the beamformer output for each of a predefined set of weights, and selects the one that best fulfils the MVDR criterion. The LCA creates images with a quality comparable to MVDR, and it is perfectly suited for a GPU.
\end{abstract}
\vspace{2cm}
(Forsiden genereres av ASA.)
% \ \\
% Total: 195 words (200 max)
% 
% \ \\
% Noen tanker jeg har selv naa:
% \begin{itemize}
% \item Det mangler noe om bruksomraader.
% \item Maa passe paa aa ikke selge LCA saa bra at MVDR ser helt haaplos ut(?)
% \end{itemize}
% 
% 
% \ \\
% Questions they wanted answered:
% \begin{itemize}
% \item This session will feature work on adapting computationally demanding modeling and signal processing tasks to take advantage of various parallel architectures, such as cloud computing, computing clusters, multicore CPUs, multi-core graphic processing units (GPUs), DSP chips, and FPGAs.
% \item If algorithm is slow - What technology will help me?
% \item What skills are needed?
% \item What tools are available and what's the cost?
% \item What acceleration factor are we looking at?
% \item How can I get started?
% \item What we did. Which tools? Was it hard? Speedup?
% \end{itemize}

\newpage
\section{Introduction}

@@

We propose three different solutions to the complexity issue depending on system size. For smaller systems we use the standard time-domain MVDR method, arithmetically optimize the estimation step, and show that image creation rates of more than 1\,Mpx/s can be achieved when this algorithm is run on a GPU. For medium sized systems we suggest applying the MVDR in the frequency domain (beamspace) instead, which allows for comparable processing rates at a negligible reduction in image quality. Finally, for large systems the estimation and inversion step can be avoided completely by utilising the Low Complexity Adaptive (LCA) beamformer.\todo{disse tre første avsnittene burde sikkert vært kortet ned mye men...}
% 
% 
% The end design is more than 2 orders of magnitude faster than a C implementation we started off with, it took considerable effort to get there.
% \\
% << MOTIVATION >>
% \\

@@

This article is outlined as follows: First we will present\todo{denne formuleringstypen begynner aa bli oppbrukt...} background material on the MVDR method operated in time and frequency domain, and on the LCA beamformer. For each method we start with the mathematical formulation, then evaluate complexity issues, and finally give our thoughts on some implementation aspects. The results are collected in \ref{implementation}. Finally, we conclude in section \ref{discssion}.
